\section{Diagnostic Analysis}

\subsection{Why Do Top Lightweight Systems Cluster?}
The tight clustering of BM25, FAISS, Naive RAG, and Simple KG (81.0--82.1\%) is a striking finding.
All four systems achieve identical L2 accuracy (100.0\%) and nearly identical L1 accuracy (80.2--81.5\%).
This suggests that for BiTempQA's short-context scenarios (2--7 memory writes per scenario), retrieval design matters less than answer generation quality---all systems retrieve sufficient context for the LLM generator to reason about bitemporal relationships.

We hypothesize that the bottleneck shifts from retrieval to generation as context length decreases below a threshold.
With only 2--7 facts per scenario, even a naive concatenation strategy (Naive RAG) retrieves the full relevant context within top-$k{=}5$.
Future work should test this hypothesis with longer scenarios (50+ writes) where retrieval selectivity becomes critical.

\subsection{Production Systems: Compression Destroys Temporal Detail}
The most striking finding is the dramatic gap between lightweight baselines (81--82\%) and production memory systems (2--8\%).
We identify three structural causes:

\paragraph{Memory compression.}
Mem0's LLM-based memory extraction (with \texttt{infer=True}) compresses raw memory writes into generalized ``memory facts.''
For example, ``Microsoft CEO changed from Ballmer to Nadella on 2014-02-04'' may be compressed into ``Satya Nadella is Microsoft CEO,'' losing the event-time/record-time distinction essential for bitemporal queries.
This explains why Mem0 achieves 19.2\% on entity attribute evolution (where some facts survive compression) but 0--1.9\% on other types.

\paragraph{Graph construction overhead.}
Graphiti builds entity-relation graphs from each memory write using LLM extraction and stores them in Neo4j.
For short scenarios (2--7 writes), the overhead of graph construction (entity resolution, relation extraction, deduplication) introduces errors that compound, leading to near-zero retrieval accuracy.
Graphiti's perfect accuracy on relationship evolution (100.0\%) vs.\ near-zero on other types suggests it captures relational changes but loses attribute-level temporal detail.

\paragraph{Temporal indexing mismatch.}
Both production systems implement their own temporal indexing, but it does not align with BiTempQA's bitemporal model.
Mem0 timestamps memories with creation time but does not preserve event time as a first-class dimension.
Graphiti uses reference time for search but lacks explicit event-time/record-time separation.

\subsection{ChromaDB's Metadata Filtering Failure}
ChromaDB's 22.6\% accuracy warrants detailed analysis.
Unlike the other lightweight systems that use embedding similarity for retrieval, ChromaDB relies on metadata filtering: temporal conditions are converted to metadata queries that filter documents before similarity ranking.
This approach fails in practice for two reasons:

\paragraph{Heterogeneous temporal formats.}
Real-world data from ChronoQA and curated events uses diverse temporal expressions. ChromaDB's metadata filtering requires consistent ISO format, and when parsing fails, queries fall back to recency-based retrieval that misses relevant context.

\paragraph{Metadata coverage gaps.}
Some memory writes lack explicit temporal metadata (e.g., relational facts without clear event times). When metadata is missing, ChromaDB cannot apply temporal filters, effectively excluding these facts from retrieval.

Notably, ChromaDB achieves perfect accuracy on L2 questions (100.0\%) but only 19.8\% on L1 questions. This suggests it can handle aggregation tasks where all relevant facts are retrieved, but fails on point-in-time queries requiring specific temporal filtering.

\subsection{Real-World vs.\ Synthetic Performance}
The merged dataset enables a natural experiment: 215 real-source scenarios vs.\ 118 LLM-generated scenarios.
While the test set is entirely real-source (by design), the training set includes both sources.
We observe that lightweight systems perform comparably on both real and synthetic scenarios in the training data, validating our LLM augmentation approach.
The rare scenario types covered only by LLM generation (contradictory information, future-dated information, entity identity resolution) remain important for benchmark completeness even though they do not appear in the test set.

\subsection{Implications for Memory System Design}
Our findings suggest four design recommendations:

\paragraph{1. Preserve raw temporal context.}
The dramatic failure of production systems shows that memory compression destroys bitemporal detail. Memory systems should preserve original timestamps and temporal context rather than compressing them into generalized facts.

\paragraph{2. Invest in temporal-aware generation for short contexts.}
For short-context scenarios, retrieval quality is not the differentiator. The LLM generator's ability to reason about bitemporal relationships from retrieved context is the primary accuracy driver.

\paragraph{3. Avoid rigid temporal metadata schemas.}
ChromaDB's failure demonstrates that metadata filtering with strict format requirements cannot handle the diversity of real-world temporal expressions.

\paragraph{4. Use real-world data for benchmark validation.}
The natural temporal dynamics in ChronoQA and curated events provide more ecologically valid test scenarios than purely synthetic templates.
