\section{Conclusion}

We presented BiTempQA, the first diagnostic benchmark explicitly designed to evaluate bitemporal reasoning (reasoning about \emph{when events occurred} vs.\ \emph{when the system learned about them}) in LLM agent memory systems.
Unlike prior benchmarks that use purely synthetic data, BiTempQA v3 is grounded in real-world temporal data: 64.6\% of scenarios come from ChronoQA (Chinese news QA) and manually curated public events, with the test set being 100\% real-source.
Our evaluation of seven memory backends (five lightweight, two production) reveals four key findings.

First, \textbf{retrieval design matters less than expected} for short-context scenarios: BM25, FAISS, Naive RAG, and Simple KG achieve statistically indistinguishable accuracy (81.0--82.1\%), suggesting that all systems retrieve sufficient context when scenarios contain fewer than 10 memory writes.

Second, \textbf{production memory systems are not yet temporally capable}: Mem0 (8.3\%) and Graphiti (2.4\%) dramatically underperform lightweight backends, primarily due to aggressive memory compression that destroys bitemporal detail and temporal indexing that does not preserve event-time/record-time separation.

Third, \textbf{rigid temporal metadata filtering fails catastrophically}: ChromaDB's 22.6\% accuracy reflects the fundamental mismatch between metadata-based temporal queries and the heterogeneous temporal formats in real-world data.

Fourth, \textbf{real-world data validates the benchmark}: the balanced gap distribution in ChronoQA-sourced scenarios and diverse domains confirm that bitemporal reasoning challenges exist in naturally-occurring data, and LLM-based data augmentation produces comparable difficulty levels.

These findings provide actionable guidance for building temporally-aware memory systems: preserve raw temporal context rather than compressing it, invest in generation quality for short contexts, avoid rigid temporal schemas, and ground benchmarks in real-world temporal dynamics.

Future work includes: human expert validation of QA pairs, English translation, expansion to longer scenarios with 50+ memory writes where retrieval selectivity matters, and evaluation of bitemporal-aware memory architectures that preserve both temporal dimensions as first-class citizens.
