\section{Experimental Setup}

\subsection{Memory Backends}
We evaluate seven memory backends spanning different retrieval paradigms.
All systems receive identical inputs: each memory write carries its text content, event time ($t_e$), and record time ($t_r$).

\textbf{FAISS}~\citep{johnson2019faiss} uses dense vector similarity search with \texttt{paraphrase-multilingual-MiniLM-L12-v2} embeddings and top-$k{=}5$ retrieval.

\textbf{BM25} uses sparse keyword matching via rank-BM25 with jieba Chinese tokenization and top-$k{=}5$ retrieval.

\textbf{Naive RAG} uses sequential passage concatenation (no relevance ranking), top-$k{=}5$.

\textbf{Simple Knowledge Graph} is an in-memory graph with entity/relation extraction and versioned attribute storage. Entity attributes maintain a history of values with valid-from/valid-until timestamps, enabling point-in-time queries. Top-$k{=}5$.

\textbf{ChromaDB}~\citep{chromadb2023} is a persistent vector database using the same \texttt{paraphrase-multilingual-MiniLM-L12-v2} embedding model with metadata filtering on temporal fields and top-$k{=}5$ retrieval.

\textbf{Mem0}~\citep{mem02025} is a production memory system that extracts structured memories via LLM (Qwen2.5-32B-Instruct) and stores them in Qdrant with semantic search. We enable its full inference pipeline (\texttt{infer=True}) for memory extraction, deduplication, and entity resolution.

\textbf{Graphiti}~\citep{graphiti2025} is Zep's temporal knowledge graph system that builds entity-relation graphs from episodes using LLM extraction and Neo4j storage. We pass temporal parameters (\texttt{reference\_time} from record time, \texttt{start\_time}/\texttt{end\_time} bounds) to its search API.

All seven systems share identical generation prompts and the answer-generation LLM backbone (DeepSeek-V3 via SiliconFlow API).
The five lightweight backends use the same embedding model, top-$k$ parameter, and chunking strategy, ensuring that performance differences reflect retrieval design.
Mem0 and Graphiti use Qwen2.5-32B-Instruct as their internal LLM for memory extraction; we discuss the implications of this difference in \S\ref{sec:limitations}.

\subsection{Evaluation Fairness}
To ensure a fair comparison, we control for the following variables:

\paragraph{Unified embedding model.}
All vector-based systems (FAISS, ChromaDB) use the same \texttt{paraphrase-multilingual-MiniLM-L12-v2} model (384 dimensions), eliminating embedding quality as a confound.

\paragraph{Unified top-$k$.}
All systems retrieve exactly $k{=}5$ context passages. Systems that internally retrieve more candidates are truncated to 5 before passing to the answer generator.

\paragraph{Record time propagation.}
All systems receive record time ($t_r$) alongside event time ($t_e$) for each memory write. Systems that support temporal metadata (ChromaDB, Simple KG, Graphiti) can use $t_r$ for filtering; others store it as auxiliary context.

\paragraph{Full feature usage.}
Mem0 runs with \texttt{infer=True} to enable its full LLM-based memory extraction pipeline. Graphiti receives temporal search parameters (\texttt{reference\_time}, \texttt{start\_time}/\texttt{end\_time}) to leverage its temporal indexing capabilities.

\paragraph{Scenario reset.}
Memory stores are reset between scenarios to prevent cross-scenario contamination.

\subsection{Statistical Methods}
We employ: (1) Bootstrap 95\% confidence intervals (10,000 iterations); (2) McNemar's test with continuity correction for pairwise system comparisons.
