\section{Related Work}

\subsection{Temporal Reasoning Benchmarks}
Recent years have seen a growing number of temporal reasoning benchmarks for LLMs.
\textsc{TimE}~\citep{time2025} provides 38,522 questions across 11 subtasks at 3 difficulty levels and evaluates LLMs' ability to understand temporal expressions and reason about event sequences.
\textsc{ETRQA}~\citep{etrqa2024} consolidates event-temporal reasoning datasets for reading comprehension.
\textsc{Test of Time}~\citep{testoftime2025} introduces synthetic temporal reasoning tasks.
\textsc{MusTQ}~\citep{mustq2024} evaluates multilingual temporal KGQA across 10 languages.
% 中文注释：补 all of them / that is，使句子更完整，不改变论点。
However, all of them treat time as a \emph{single} dimension: none systematically distinguish event occurrence time from information recording time, a distinction that is critical for memory systems.

\subsection{Memory System Benchmarks}
% 中文注释：补 the development of，更符合正式学术写作习惯。
The emergence of memory-augmented LLM agents has also driven the development of dedicated memory benchmarks.
LoCoMo~\citep{locomo2024} evaluates long conversational memory with 1,986 questions across 5 types, including adversarial and temporal categories.
LongMemEval~\citep{longmemeval2024} provides 4,500 questions for evaluating 8 memory systems on single-hop, multi-hop, and temporal reasoning.
MemoryAgentBench~\citep{memoryagentbench2025} focuses on memory evaluation across multiple datasets.
While these benchmarks include temporal questions, they do not systematically test \emph{dual-timestamp} reasoning, namely the ability to distinguish ``when this happened'' from ``when I learned about it.''

\subsection{Bitemporal Data Modeling}
The distinction between event time (valid time) and record time (transaction time) is foundational in temporal database research~\citep{snodgrass1999tsql2,jensen1999temporal}.
% 中文注释：like 改为 such as，并统一引号内首字母大写，只做英文规范化。
Bitemporal data models track both dimensions, enabling queries such as ``What did the system know at time $T_1$ about events that occurred at time $T_2$?''
This concept has been extensively studied in database systems, but it has rarely been brought into the evaluation of LLM memory systems.
BiTempQA extends this line of thinking to memory-system evaluation by making bitemporal reasoning an explicit target of assessment.

\subsection{Positioning}
Table~\ref{tab:benchmark_comparison} positions BiTempQA relative to existing work.
To the best of our knowledge, BiTempQA is currently the only benchmark that both evaluates dual-timestamp reasoning and tests multiple memory-system backends.

\begin{table}[t]
\centering
% 中文注释：这张表列数较多，恢复按栏宽缩放，避免双栏模板下横向溢出。
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccccc}
\toprule
\textbf{Benchmark} & \textbf{Year} & \textbf{Temp.} & \textbf{Dual TS} & \textbf{Mem} & \textbf{Lang} & \textbf{Scale} \\
\midrule
LoCoMo & 2024 & Partial & \xmark & \cmark & EN & 1,986 \\
LongMemEval & 2024 & \xmark & \xmark & \cmark & EN & 4,500 \\
TimE & 2025 & \cmark & \xmark & \xmark & EN & 38,522 \\
ETRQA & 2025 & \cmark & \xmark & \xmark & EN & Consol. \\
Test of Time & 2025 & \cmark & \xmark & \xmark & EN & Novel \\
MusTQ & 2024 & \cmark & \xmark & \xmark & Multi & 15K \\
\midrule
\textbf{BiTempQA} & \textbf{2026} & \cmark & \cmark & \cmark & ZH & 1,536 \\
\bottomrule
\end{tabular}%
}
\caption{Comparison with related benchmarks. \textbf{Dual TS}: evaluates dual-timestamp reasoning. \textbf{Mem}: tests memory backends (not just LLMs).}
\label{tab:benchmark_comparison}
\end{table}
