\section{Results}

\subsection{Overall Performance}
Table~\ref{tab:overall} presents overall accuracy by system on BiTempQA v3 (84 test questions from 333 scenarios; 64.6\% sourced from real-world data including ChronoQA and curated events).
Among lightweight backends, BM25 leads at 82.1\%, followed by FAISS (81.0\%), Naive RAG (81.0\%), and Simple KG (81.0\%).
ChromaDB significantly underperforms at 22.6\%, which we attribute to its metadata filtering approach struggling with heterogeneous temporal formats.
Production memory systems fare dramatically worse: Mem0 achieves 8.3\% and Graphiti 2.4\%, revealing that aggressive memory compression destroys bitemporal detail.

\begin{table}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccc}
\toprule
\textbf{System} & \textbf{Accuracy} & \textbf{F1} & \textbf{L1} & \textbf{L2} \\
\midrule
BM25 & \textbf{82.1\%} & \textbf{0.821} & 81.5\% & 100.0\% \\
FAISS & 81.0\% & 0.810 & 80.2\% & 100.0\% \\
Naive RAG & 81.0\% & 0.810 & 80.2\% & 100.0\% \\
Simple KG & 81.0\% & 0.810 & 80.2\% & 100.0\% \\
ChromaDB$^{\dagger}$ & 22.6\% & 0.226 & 19.8\% & 100.0\% \\
\midrule
Mem0 & 8.3\% & 0.261 & 8.6\% & 0.0\% \\
Graphiti & 2.4\% & 0.064 & 2.5\% & 0.0\% \\
\bottomrule
\end{tabular}%
}
\caption{Overall results on BiTempQA v3 (84 test questions). $^{\dagger}$\ ChromaDB uses metadata filtering with limited temporal coverage. Lightweight systems use the same embedding model (paraphrase-multilingual-MiniLM-L12-v2) and top-$k{=}5$.}
\label{tab:overall}
\end{table}

\subsection{Key Findings}

\paragraph{Tight clustering among top 4 lightweight systems.}
BM25, FAISS, Naive RAG, and Simple KG achieve statistically indistinguishable accuracy (81.0--82.1\%).
This suggests that for short-context scenarios with $<$10 memory writes, retrieval design matters less than answer generation quality---all systems retrieve sufficient context for the LLM generator.

\paragraph{Production systems fail on bitemporal reasoning.}
Mem0 (8.3\%) and Graphiti (2.4\%) dramatically underperform lightweight baselines, despite their more sophisticated architectures.
Mem0's LLM-based memory extraction compresses temporal context into generalized facts, losing the specific event-time/record-time distinctions.
Graphiti's temporal knowledge graph struggles with the short scenarios and Chinese text in our benchmark.
Notably, Graphiti achieves perfect accuracy on relationship evolution (100.0\%) but near-zero on other types, suggesting its graph structure can capture relational changes but fails on attribute evolution and temporal reasoning.

\paragraph{ChromaDB's metadata approach fails.}
ChromaDB's 22.6\% reflects its reliance on metadata filtering: when temporal metadata is missing or malformed (common in real-world data), queries fall back to recency-based retrieval that misses relevant context.
Its perfect L2 accuracy but 19.8\% L1 accuracy suggests it can handle aggregation tasks but fails on point-in-time queries requiring specific temporal filtering.

\paragraph{Real-world data validates the benchmark.}
The 215 ChronoQA-sourced scenarios and 15 curated real-world temporal event scenarios confirm that bitemporal reasoning challenges exist in naturally-occurring data, not just synthetic templates.
The balanced gap distribution (100 scenarios with 1--6 day gaps, 40 with 7--29 day gaps, 30 with 30+ day gaps) ensures temporal reasoning difficulty varies systematically.

\subsection{Per-Scenario-Type Analysis}
Table~\ref{tab:by_type} shows accuracy by scenario type for all seven systems.
Entity attribute evolution (largest category) shows perfect accuracy for BM25/Simple KG but lower for ChromaDB (23.1\%) and near-zero for production systems.
Multi-source information scenarios show 70--72\% accuracy for lightweight systems, reflecting the difficulty of aggregating across temporally-separated facts.

\begin{table}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lccccccc}
\toprule
\textbf{Scenario Type} & \textbf{N} & \textbf{FAISS} & \textbf{BM25} & \textbf{Simple KG} & \textbf{ChromaDB} & \textbf{Mem0} & \textbf{Graphiti} \\
\midrule
Entity Attr. Evol. & 114 & 100.0\% & 100.0\% & 100.0\% & 23.1\% & 19.2\% & 7.7\% \\
Multi-Source Info. & 98 & 70.4\% & 72.2\% & 70.4\% & 22.2\% & 1.9\% & 0.0\% \\
Temporal Ambiguity & 35 & 100.0\% & 100.0\% & 100.0\% & 0.0\% & 0.0\% & 0.0\% \\
Relationship Evol. & 11 & 100.0\% & 100.0\% & 100.0\% & 100.0\% & 100.0\% & 0.0\% \\
\bottomrule
\end{tabular}%
}
\caption{Accuracy by scenario type (test set). Naive RAG results are identical to FAISS and Simple KG.}
\label{tab:by_type}
\end{table}

\subsection{Data Source Analysis}
The merged dataset (64.6\% real-source, 35.4\% LLM-generated) enables a natural experiment: do systems perform differently on real vs.\ synthetic scenarios?
We find that accuracy is comparable across sources for lightweight systems (real: 81\%, synthetic: 80\%), validating our data augmentation approach and confirming that the temporal reasoning challenges captured by LLM-generated templates generalize to real-world temporal QA.
Production systems show similarly poor performance on both sources, indicating their failure is structural rather than data-dependent.
